---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit the README.Rmd file -->


```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Lab report #1

Follow the instructions posted at https://ds202-at-isu.github.io/labs.html for the lab assignment. The work is meant to be finished during the lab time, but you have time until Monday evening to polish things.

Include your answers in this document (Rmd file). Make sure that it knits properly (into the md file). Upload both the Rmd and the md file to your repository.  

All submissions to the github repo will be automatically uploaded for grading once the due date is passed.
Submit a link to your repository on Canvas (only one submission per team) to signal to the instructors that you are done with your submission.

```{r}
library(classdata)
library(dplyr)
library(ggplot2)

data(ames)
summary(ames)
head(ames$'Parcel ID')
typeof(ames$'Parcel ID')
str(ames)
head(ames$'Sale Price')
range(ames$'Sale Price', na.rm=TRUE)
```

1. Inspecting the dataset
Manas:
Parcel ID (Character): String containing a set of numbers to describe a unique ID for each parcel. The string can be a 10 digit ID value beginning with a 0.

Address (Character): String containing the address to which the parcel is addressed to. All address strings are in the form of "Number Street, Ames". They do not have a definite data range as they're strings, but they'll all be within Ames.

Style (Factor): This is a factor variable with 12 levels of strings which describes a different style layout of the house. It's range is between the 12 possible levels.

Occupancy (Factor): This is a factor variable with 4 different levels of strings which describes the amount of people that can live in a home. The ranges are between Condominium Single-Family / Owner Occupied, Townhouse, Two-Family Conversion Two-Family Duplex

Sale Date (Date): This date variable describes the day the house was sold. The range of this variable is between 2017-07-03 and 2022-08-31

Sale Price (Numeric): This variable describes the amount the house was sold in USD. The range of this variable is between 0 and 20,500,000.


Multi Sale (Character): Indicates whether the parcel has multiple sales associated with it. This variable contains "Yes" or "No" values, though more detailed levels would need further investigation.

Year Built (Numeric): Range: From 0 (which seems like an error or placeholder) to 2022. Filter out invalid years if needed.
Description: The range of years when properties were built spans from 1880 (after filtering invalid data) to 2022.

Acres (Numeric): Range: From 0.0000 to 12.0120 acres.
Description: Lot sizes range from small parcels (0 acres) to large estates (12 acres).

Total Living Area (Numeric): Range: From 0 (possible data issue or very small homes) to 6007 square feet.
Description: The living area ranges from 0 to 6007 square feet, with most properties falling between 1095 and 1792 square feet.

Bedrooms (Numeric): Range: From 0 (no bedroom) to 10 bedrooms.
Description: The number of bedrooms ranges from 0 (indicating possibly non-residential or studio properties) to 10, with most homes having 3 to 4 bedrooms.

Finished Basement Area (Numeric): Range: From 10 square feet (indicating very little finished area) to 6496 square feet.
Description: Finished basement areas range from 10 to 6496 square feet, with most properties having between 474 and 1011 square feet.

Lot Area (Numeric): Range: From 0 square feet (possibly indicating missing data) to 523,228 square feet.
Description: Lot sizes vary significantly, from 0 to 523,228 square feet, with typical lot sizes ranging from 6553 to 12,088 square feet.

AC (Character): Description: This variable likely indicates whether air conditioning is present or absent (e.g., "Yes"/"No" or other similar categories), but the summary table does not show specific values. You'll need to check the unique values for further clarification.

FirePlace (Character): Description: Indicates the presence or absence of a fireplace. Similar to the "AC" variable, the summary does not provide specific values, but you could describe it based on whether fireplaces are present.

Neighborhood (Factor): Description: A factor variable with multiple levels indicating different neighborhoods in Ames. The summary shows some of the largest neighborhoods like North Ames (854 observations) and College Creek (652 observations).

2. Our main variable of special interest in this dataset is the sale price of the homes


3. 
```{r}
new_data <- filter(ames, `Sale Price` < 10000000, `Sale Price` > 0)
ggplot(data=ames, aes(x=`Sale Price`)) + geom_histogram(binwidth=8000)

mean(new_data$`Sale Price`)
```
Since our main variable is numeric, we chose a histogram. Taking a dive into the chart, we can see that the mean price of a home is at \$1,017,479. Majority of the homes are less than \$1,000,000, the mean is skewed due to the few houses which are outliers that are extremely expensive. Taking the outliers > \$1,000,000 out of the equation, the average home is \$241,960.


4. 
Manas:
```{r}
new_data <- filter(ames, `YearBuilt` > 1500, `Sale Price` > 0, !is.na(`YearBuilt`), !is.na(`Sale Price`), `Sale Price` < 1000000)
ggplot(data=new_data, aes(x=`YearBuilt`, y=`Sale Price`)) + geom_point()
range(ames$YearBuilt)
```
I picked the YearBuilt variable to see if there's a realationship with the main variable Sale Price. overall range of YearBuilt is 0 to 2022. This data needed some filtering, so the real min data shows the first houses in the dataset are from 1880. I also removed houses that are well over \$1M, as they're houses that which are outliers to the dataset. The mean shows us that the average house was built in 1976. Using a scatterplot, we can see a positive correlation between the YearBuilt and the sales price, as the average house price has continued to rise as the years go on.

Muhammad:
```{r}
living_area_data <- filter(ames, `TotalLivingArea (sf)` > 0, `Sale Price` > 0, `Sale Price` < 1000000)
ggplot(data = living_area_data, aes(x = `TotalLivingArea (sf)`, y = `Sale Price`)) + geom_point() + geom_smooth(method = "lm", col = "blue")
range(living_area_data$`TotalLivingArea (sf)`)

```
Analysis:

Total Living Area Range: From 334 to 5642 square feet.

Correlation: A positive correlation is visible, as larger living areas tend to have higher sale prices. The scatter plot with a trend line shows that as living area increases, the sale price generally rises.

Observations: Larger living spaces contribute significantly to higher property values, confirming that property size is a strong indicator of sale price.

Shivansh:
```{r}
acres_bedrooms_data <- filter(ames, Acres > 0, Bedrooms > 0)

ggplot(data = acres_bedrooms_data, aes(x = Acres, y = Bedrooms)) +
  geom_point() +
  geom_smooth(method = "lm", col = "blue")
  labs(title = "Relationship between Acres and Bedrooms",
       x = "Acres",
       y = "Number of Bedrooms")

range(acres_bedrooms_data$Acres)
range(acres_bedrooms_data$Bedrooms)
```

Analysis:

"Acres" and "Bedrooms" show a positive correlation, but the relationship is not extremely strong due to the wide variance at lower acre sizes. While larger lots generally have more bedrooms, the size of the lot is not the sole determinant of the number of bedrooms, especially for properties with less than 1 acre. The property’s style (e.g., single-family home versus multi-family home) or the neighborhood, might influence the number of bedrooms independent of acreage.
